### 2024-01-28 Findings (Shared Folder)/Chapters/MeetingLog.tex

\chapter{Meeting log}

\subsection{29 Jan Meeting with Leah}
Things to discuss
\begin{itemize}
	\item Citation with Year-Author style, or Number Style?
\end{itemize}