### 2024-01-28 Findings (Shared Folder)/Findings.tex

%\documentclass[12pt,a4paper]{book}
%\usepackage{commands}
%
%\begin{document}
%
%	
%\end{document}
\documentclass{labdiary}
\usepackage{commands}
\usepackage{natbib} 
%\setcitestyle{authoryear, open={([},close={)]}}
%\setcitestyle{numbers,square}
\bibliographystyle{dinat}

\title{My Findings}
\author{Ali Fele Paranj}
\date{\today}

\begin{document}

	\maketitle

	\tableofcontents

%	\include{Chapters/StatisticalPhysicsForBiology.tex}
	\include{Chapters/modelAttempts.tex}
	\include{Chapters/PapersReviewed.tex}
	\include{Chapters/MolecularBiology.tex}
	\include{Chapters/Comments.tex}
	\include{Chapters/MeetingLog.tex}

	% Bibliography
%	\bibliographystyle{plainnat}

	\bibliography{References/references} % The filename of the .bib file (without the extension)

\end{document}