### 2024-01-28 Findings (Shared Folder)/Chapters/modelAttempts.tex
new comments added

\chapter{Modeling Attempts}

\section{Simple ODE model (First Iteration)}
Here, we develop a model that keeps track of the following variables
\begin{quote}
	$n(t)$= density of tip cells in area of interest, (number per unit area).

	$\rho(t)$ = density of blood vessels (length per unit area).

	$c(t)$ = concentration of drug delivered to region by blood vessels.
\end{quote}
An updating list of model parameters
\begin{itemize}
	\item $v$ [length/time]: The rate at which the tip cells move and extends the blood vessels.
	\item $\delta_v$ [1/time]: The rate at which the vascular structure gets degraded.
	\item $\lambda_s$ [1/time]: Tip cell division rate (splitting rate).
	\item $\lambda_b$ [1/time/length]: Tip cell emerging rate from stalk cells.
	\item $\delta_t$ [1/time]: Tip cell death/deactivation rate.
	\item $\kappa$ [area/length/time]: Re-connection of tip cells to the other capillaries to form loops.  
\end{itemize}

\subsubsection*{Studying dynamics of vessel formation}
\[ \frac{d\rho}{dt} = ??. \]
The active tip cells extend the vascular structure as they move. Assuming the tip cells move at rate $v$, then 
\[ \frac{d \rho}{dt} = vn+ ??. \]
Also, assuming the vascular structure degrades with rate $\delta$ [per unit time], we can add the degradation term 
\[ \boxed{\frac{d\rho}{dt} = vn + \delta_v \rho} . \]

\subsubsection*{Studying the Dynamics of Tip Cells}
Things important in the dynamics of the tip cells
\begin{itemize}
	\item Generation of the tip cells: There are at least two ways for new tip cell generation listed as follows:
	\begin{enumerate}[(i)]
		\item Splitting mechanics: When the tip cells splits new vascular stem gets two heads. This should be proportional to the density of tip cells. The parameter $\lambda_s$ [per unit time] reflects this mechanism.
		\item Branching: New tip cells can form out the the endothelial stalk cells. This process should be proportional to the density of blood vessels. The parameter $\lambda_b$ [per unit time per unit length] reflects this mechanism
	\end{enumerate}
	\item Loss of tip cells
	\begin{enumerate}[(i)]
		\item Death of the tip cells or getting deactivated: Reflected by the parameter $\delta_t$
		\item Joining the other branches of vascular network: When a tip cell reconnects another capillary branch, then they disappear. The parameter $\kappa$ is for this mechanism. Note that the re-connection term is proportional to both number of tips cells, as well as the density of blood vessels. Thus the units of $\kappa$ should be [area/length/time]. 
	\end{enumerate}
	\item The movement of tip cells and formation of new vascular networks along the way.
\end{itemize}

\[ \boxed{\frac{dn}{dt} = (\lambda_s - \delta_t) n + \lambda_b \rho + \kappa n \rho}.  \]

\subsubsection*{Nondimensionalization}
In order the analyze the model more easily, we nondimensionalize the system with the following change of variable
\[ \rho = R \tilde{\rho}, \qquad n = N \tilde{n}, \qquad t = T \tau. \]
There are many possible choice to choose the scaling factors $R, N, T$. However, we will choose them in a way that they are always positive, and the system of ODE becomes as simple as possible. Substituting the change of variable above in the ODE system, we will get
\begin{align*}
	&\frac{d\tilde{\rho}}{d\tau} = \frac{vNT}{R}\tilde{n} + \delta_v T \tilde{\rho},\\
	&\frac{d\tilde{n}}{d\tau} = T(\lambda_s-\delta_t)\tilde{n} + \frac{\lambda_b T R}{N} \tilde{\rho} + T\kappa R \tilde{n}\tilde{\rho}.
\end{align*}
We choose the following values for $T,N$, and $R$
\[ T = \frac{1}{\delta_v}, \qquad R = \frac{\delta_v}{\kappa}, \qquad N = \frac{\lambda_b}{\kappa}. \]
This is a very suitable moment to pause and check the dimensions if they match (I did it and all of them matches!). With these choices from the coefficients, the system of ODEs will be
\[ \frac{d\tilde{n}}{d \tau} = \frac{\lambda_s - \delta_t}{\delta_v} \tilde{n} + \tilde{\rho} + \tilde{n}\tilde{\rho}, \qquad
\frac{d\tilde{\rho}}{d\tau} = \frac{v\lambda_b}{\delta_v^2}\tilde{n} + \tilde{\rho}. \]
To make the ODEs simpler to work with, we will write $n,\rho$ in place of $\tilde{n}$ and $\tilde{\rho}$, and also we introduce the following parameters
\[ \alpha = \lambda_s - \delta_t, \qquad \beta =\delta_v, \qquad \gamma = v\lambda_b. \tag{$\clubsuit$} \]
Then we can write
\begin{equation*}
	\boxed{
		\begin{aligned}
			&\dot{n} = \frac{\alpha}{\beta}n + \rho + n\rho, \\
			&\dot{\rho} = \frac{\gamma}{\beta^2}n +\rho.
		\end{aligned}
	}
	\tag{\smiley}
\end{equation*}
After some algebra, it turns out that there are two equilibrium points for this system.
\[ p^0_1 = (0,0), \qquad p^0_2 = (\frac{\alpha\beta}{\gamma} - 1, \frac{\gamma}{\beta^2}-\frac{\alpha}{\beta}) = (\frac{\alpha\beta-\gamma}{\gamma},\frac{\gamma - \alpha\beta}{\beta^2}). \]
\begin{strangeObs}
	Something does not make sense! We know that the variables $n$ and $\rho$ must all be positive, as they are both representing biological values. However, by looking closely at the second component of $p^0_2$, it turns out th at $\gamma - \alpha\beta$ should be positive. In order to have the first component positive as well, then $\gamma$ should be negative, which is not possible. That is because by $(\clubsuit)$ we know that $\gamma>0$. There is at least one solution for this paradox, which is letting $\alpha\beta - \gamma =0$. Then this equilibrium point will be the same as $p^0_1$, showing the fact that the system has only one equilibrium point with biological meaning, which is the origin (not very interesting)! \\
	However, I am going to ignore this observation for now and proceed with the stability analysis.

	\label{strangeObs:SimpleODEequiliNegative}
\end{strangeObs}

In order to analyze the stability of these equilibrium points, we first need to calculate the Jacobian matrix of the ODE system 
\[ DF = \matt{\alpha/\beta+\rho}{1+n}{\gamma/\beta^2}{1}. \]

\subsubsection{Stability Analysis of $p^0_2$}
By evaluating the Jacobian matrix at the equilibrium point we will have
\[ DF[p^0_2] = \matt{\gamma/\beta^2}{\alpha\beta/\gamma}{\gamma/\beta^2}{1}. \]
The trace and determinant of this matrix is
\[  \Delta = \gamma/\beta_2 - \alpha/\beta, \qquad \sigma=\gamma/\beta^2 + 1.  \]
By close inspection, it turns out that $\Delta$ is the same as the first component of $p^0_2$, which should be positive. This implies $\Delta > 0$. So the sign of the trace of the Jacobian matrix will determine the stability. From $(\clubsuit)$, $\gamma>0$. Thus $\sigma > 0$. This indicates that the equilibrium point $p_0^2$ is unstable (look at \autoref{strangeObs:SimpleODEequiliNegative} for another ongoing issue regarding this equilibrium point).

\subsubsection{Stability Analysis of $p^0_1$}
The Jacobian matrix evaluated at $p^0_1$ is
\[ DF[p^0_1] = \matt{\alpha/\beta}{1}{\gamma/\beta^2}{1}. \]
The determinant and the trace will be
\[  \Delta = \alpha/\beta - \gamma/\beta^2, \qquad \sigma = \alpha/\beta+1. \]
Since the determinant is minus the first component of $p^0_2$. Thus $\delta<0$. This implies that regardless of the value of $\sigma$, the equilibrium $p^0_1$ is a saddle point.

\begin{strangeObs}
	This equilibrium point $p^0_1$ has no choice but to be stable, if we expect the model to represent real biological situation. But our analysis above, reveals that this equilibrium point is in fact a saddle point. However, I can argue that if that the eigenvector with negative eigenvalue for the Jacobian matrix evaluated at $p^0_2$, lies in the first quadrant, then $p^0_2$ will ``look'' stable for the points at the first quadrant. But I am not sure if this is something normal to happen for a biological model.
\end{strangeObs}

\newpage

\section{Some Ideas to Try}
This section might have very simple, basic and sometimes silly ideas that came into my mind during developing some models and I thought they might worth trying
\begin{itemize}
	\item Developing a model for a weighted graph generation. I suspect a weighted graph might have all the necessary information we want.
\end{itemize}