### 2024-06-09 Findings (Shared Folder)/images/orbitApproximation.tex

\begin{figure}[h!]
	\centering

	\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

	\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
		%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

		%Straight Lines [id:da7253651408536048] 
		\draw [color={rgb, 255:red, 160; green, 160; blue, 160 }  ,draw opacity=1 ][line width=0.75]    (136.83,103.17) -- (149.81,85.61) ;
		\draw [shift={(151,84)}, rotate = 126.47] [color={rgb, 255:red, 160; green, 160; blue, 160 }  ,draw opacity=1 ][line width=0.75]    (6.56,-1.97) .. controls (4.17,-0.84) and (1.99,-0.18) .. (0,0) .. controls (1.99,0.18) and (4.17,0.84) .. (6.56,1.97)   ;
		%Curve Lines [id:da4933662152703622] 
		\draw    (82,133.33) .. controls (151.8,130.12) and (147.11,57.99) .. (164.66,37.97) ;
		\draw [shift={(166.67,36)}, rotate = 140.19] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (7.14,-3.43) -- (0,0) -- (7.14,3.43) -- (4.74,0) -- cycle    ;
		%Curve Lines [id:da8486116401578401] 
		\draw    (164,38.33) .. controls (173,29.67) and (176.33,28.67) .. (188,22.33) ;
		%Shape: Circle [id:dp6250392978342221] 
		\draw  [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ,fill opacity=1 ] (135,103.17) .. controls (135,102.15) and (135.82,101.33) .. (136.83,101.33) .. controls (137.85,101.33) and (138.67,102.15) .. (138.67,103.17) .. controls (138.67,104.18) and (137.85,105) .. (136.83,105) .. controls (135.82,105) and (135,104.18) .. (135,103.17) -- cycle ;
		%Shape: Circle [id:dp5689260707692063] 
		\draw  [fill={rgb, 255:red, 74; green, 144; blue, 226 }  ,fill opacity=1 ] (149.33,70.5) .. controls (149.33,69.49) and (150.15,68.67) .. (151.17,68.67) .. controls (152.18,68.67) and (153,69.49) .. (153,70.5) .. controls (153,71.51) and (152.18,72.33) .. (151.17,72.33) .. controls (150.15,72.33) and (149.33,71.51) .. (149.33,70.5) -- cycle ;

		% Text Node
		\draw (134,107.4) node [anchor=north west][inner sep=0.75pt]  [font=\tiny]  {$X_{0}$};
		% Text Node
		\draw (151.67,89.73) node [anchor=north west][inner sep=0.75pt]  [font=\tiny,color={rgb, 255:red, 116; green, 116; blue, 116 }  ,opacity=1 ]  {$\vec{v}( X_{0})$};
		% Text Node
		\draw (156.67,61.4) node [anchor=north west][inner sep=0.75pt]  [font=\tiny,color={rgb, 255:red, 0; green, 61; blue, 133 }  ,opacity=1 ]  {$g^{\Delta t} X_{0} =X_{0} +V( X_{0}) \Delta t+O\left( \Delta t^{2}\right)$};

	\end{tikzpicture}
\end{figure}