### 2024-08-11 Findings (Shared Folder)/Chapters/branchingMorphogenesis.tex
new notes added

\chapter{Branching Morphogenesis}

\begin{itemize}[nosep]
	\item Fisher-KPP model for branching morphogenesis in VisualPDE: \href{https://visualpde.com/sim/?options=N4IgxiBcIG4JYCc4BM4GcQBoTIKq6hABEscA1MwgBlOQHU7raNoBGKj0sA6G7Aa0JEABAF5hrAHRVhcAHbCA2q0zsAugG4EAfQBGY4VWntZCxVVWa52meIDM000osBWTToCmBo6xdPzlhoANt7SMvLOmAAsmgC2oRz+FjEapLEAhtSSdgDsaXBM2AAOhADCAK5oAC4A9rGkRQg1uh4AYuVyhOUNTS0AMh5yAOZVABZQdhwWIAg8M3oAVOULABSsALQr5QDUMACUAPQr1lR7e8LrwrFLpAiU0J5LwtvCKzq6Byd7S6s7+xdBBYwW6MXikNAlaB2UjAyCKUDpBBNADupRqQRq5QQUAAbC4XHYXNhESiiIM0HAqgBPKBGIkgEk1ZEDYZjACy6QAHoRWKRGciAMpgdJBDyEdLlWp8pFMgAahD62k5BwATBdhNziTLkQBNBXaKmq9U04mSmoCjxVNEYrEAJXSwzFkAAZiK0B5sMLYh4EOkACqjS2ZSCTT3pb2+gAKowKIemXp96QAWjU6rTJAAOT3ozEIDKQ2CIFDoLg5rG6RFQV1Bd3ZuS1LEsau18A1eu56252l1hsIACiRQpGM6kCMVBUrfbWIAcuV6pB6WBKrVYgKsa6wE7mx6QB5Yroamgm26d3uD0eAIL7uCDKrp+lnw9oIhwZ3OypOow47CPo-R2NSC4ORUFmu77k+AoxnINKQOwP7gUeAqxKmYxyB4R7piq8HnmgApFB4S5BJWo6SFhYE4QGQbpjkGYuFE2DOkEcBFPhyCdli+ZVie2BMWhdAoOy5RBFAE4ZJy7F3CK5ROiARi5PkcgSWQUkyXwIA1DAPpETB27YBpWnpFSEndupmkINpA5Dm26YcPS+nmYZfaco0PKkPZ2l9PIHj8cggnCbBxQYlUfpUvhhBFERaE9M08hDFxNY7o0zRtB0XTRf0gwjOMcYcMUvQeCFYXQPIVQeEMvrCXlyXymCVUtHqtUgGg67pJu7QjrJpBjAgHgeEQAASHhwEMoxVEKIqfpIOTYJpYC1AgrQ3kEyDxS2yKjOkwU1JGQVpdgcjhjJvIAL6YAi2rGZAeIEvS-JknIFLUveWooiyWUctybDSii42iuKZrfXK+rKmqlyagy2oNSAiqGqDwgmgyZoWlaZYIPajqrTuCa+pRVTBqG4CHVGMYTPGRPJqm85GKBYCo5x0DwEgqAYNmNoIBW2Iutxk69seCU9h2qMmbTU79oOcDDjZE4i72s7zouy51GuCAblu3O-nzLYa1eug3vWz3kU+L5vh+6bfobf4k7BkhASB2EQVBMFwRbuHITUqHoSwRhkRreEEUJxHe-bR648GRg0XRDFMSxHhsXT6SQrpIC8d5AmjGyQkidgYlKSpWTydn8i50E0mFKZBk6dz7mGZd0zV1SlkS9ZJG2XpZkWc5nMgLybcV55fFpxn-kThF7uFTJEUOmKdWxZjdUpR13Tz29Yyk7lIBJS04+ECVZUVelHg1Z189Q2pzUq61C9l91vUDUNI1jcKf0kdNsAEfNi0eMtc8gOtm1+ttXaDNSAHW9FAFUx01DYGRCtaAS9f4FngRSWIfpKTPxAItNAgYEDrAANKRkjMINkNRkBf2EM6GoCBhAACFfRyDAFBIYRDKFFFGDUIYgx0IlmOkAA}{Link}.
\end{itemize}