### 2024-08-12 Findings (Shared Folder)/Chapters/branchingMorphogenesis.tex
new notes added

\chapter{Branching Morphogenesis}

\begin{itemize}[nosep]
	\item Fisher-KPP model for branching morphogenesis in VisualPDE: \href{https://visualpde.com/sim/?options=N4IgxiBcIG4JYCc4BM4GcQBoTIKq6hABEscA1MwgBlOQHU7raNoBGKj0sA6G7Aa0JEABAF5hrAHRVhcAHbCA2q0zsAugG4EAfQBGY4VWntZCxVVWa52meIDM000osBWTToCmBo6xdPzlhoANt7SMvLOmAAsmgC2oRz+FjEapLEAhtSSdgDsaXBM2AAOhADCAK5oAC4A9rGkRQg1uh4AYuVyhOUNTS0AMh5yAOZVABZQdhwWIAg8M3oAVOULABSsALQr5QDUMACUAPQr1lR7e8LrwrFLpAiU0J5LwtvCKzq6Byd7S6s7+xdBBYwW6MXikNAlaB2UjAyCKUDpBBNADupRqQRq5QQUAAbC4XHYXNhESiiIM0HAqgBPKBGIkgEk1ZEDYZjACy6QAHoRWKRGciAMpgdJBDyEdLlWp8pFMgAahD62k5BwATBdhNziTLkQBNBXaKmq9U04mSmoCjxVNEYrEAJXSwzFkAAZiK0B5sMLYh4EOkACqjS2ZSCTT3pb2+gAKowKIemXp96QAWjU6rTJAAOT3ozEIDKQ2CIFDoLg5rG6RFQV1Bd3ZuS1LEsau18A1eu56252l1hsIACiRQpGM6kCMVBUrfbWIAcuV6pB6WBKrVYgKsa6wE7mx6QB5Yroamgm26d3uD0eAIL7uCDKrp+lnw9oIhwZ3OypOow47CPo-R2NSC4ORUFmu77k+AoxnINKQOwP7gUeAqxKmYxyB4R7piq8HnmgApFB4S5BJWo6SFhYE4QGQbpjkGYuFE2DOkEcBFPhyCdli+ZVie2BMWhdAoOy5RBFAE4ZJy7F3CK5ROiARi5PkcgSWQUkyXwIA1DAPpETB27YBpWnpFSEndupmkINpA5Dm26YcPS+nmYZfaco0PKkPZ2l9PIHj8cggnCbBxQYlUfpUvhhBFERaE9M08hDFxNY7o0zRtB0XTRf0gwjOMcYcMUvQeCFYXQPIVQeEMvrCXlyXymCVUtHqtUgGg67pJu7QjrJpBjAgHgeEQAASHhwEMoxVEKIqfpIOTYJpYC1AgrQ3kEyDxS2yKjOkwU1JGQVpdgcjhjJvIAL6YAi2rGZAeIEvS-JknIFLUveWooiyWUctybDSii42iuKZrfXK+rKmqlyagy2oNSAiqGqDwgmgyZoWlaZYIPajqrTuCa+pRVTBqG4CHVGMYTPGRPJqm85GKBYCo5x0DwEgqAYNmNoIBW2Iutxk69seCU9h2qMmbTU79oOcDDjZE4i72s7zouy51GuCAblu3O-nzLYa1eug3vWz3kU+L5vh+6bfobf4k7BkhASB2EQVBMFwRbuHITUqHoSwRhkRreEEUJxHe-bR648GRg0XRDFMSxHhsXT6SQrpIC8d5AmjGyQkidgYlKSpWTydn8i50E0mFKZBk6dz7mGZd0zV1SlkS9ZJG2XpZkWc5nMgLybcV55fFpxn-kThF7uFTJEUOmKdWxZjdUpR13Tz29Yyk7lIBJS04+ECVZUVelHg1Z189Q2pzUq61C9l91vUDUNI1jcKf0kdNsAEfNi0eMtc8gOtm1+ttXaDNSAHW9FAFUx01DYGRCtaAS9f4FngRSWIfpKTPxAItNAgYEDrAANKRkjMINkNRkBf2EM6GoCBhAACFfRyDAFBIYRDKFFFGDUIYgx0IlmOkAA}{Link}.
	\item Guo pattern formation model for side branching: \href{https://visualpde.com/sim/?options=N4IgRgTiBcIAwDoCsIA0IDGMQDcCWEeAJngM5ohECqV2AIgPoCCFRAam-QwJKsDqfLgGVWAFxiI4ARnQZasAEpMAcnQBU3BkLVS4FDJ1hSKAay5MABAF4LkgEwW8AOwsBtOKikBdANyNu1rYIdgBsji7unr6MQoGSYc5uHt4+GHEIcHAOiZEpALYArulSCRHJvhAAFgD2zOlwAMzhSVE+TkU2kgAszbkVNTz1mY295akMcPXZZa0A5gCGeXnzU6OtAKYADqR4ADbVLp0IABxIaynrDFLFkzljREM9d60AZoG6574Uy9iIDQDs3zwvwoTmwXQom2wAGECqRRNU8pCINUwOsAGIFMGwFjoTYotEAGXWTlmokqMAaw3QEHkmDUTAAenY1EIAPTcAC0hQZAGoqrUmGoAJoUCCGelMlmxTkWdoaCz8gbcEViwSwDATCyyhZLeas7UWLY7fZOEWssUARWwRAZhsuUjUFmFiosAAphcyAJRst1SXkvEXeiikbEgSwBWKi9CkACOMFEEAK63QOBgrlA8wgKIA7tDqvsClBoCEkEgGkh0Fnc3QSTtRABPCTIKvZ6o54mk8kAWXmAA9sMZW7mhBh5rt1th5gUERRq+2ABrYQkMPtshyygfD9ui2Arhvrw1Nqsz6pCdaifOFiAKeakyfQF7j0gpzCLdYQeYAFUqF-mlI8N88g-eYAAVKmBaAqVkd9PwALWqRFm2OWQC2qItlihWB8EIEhyFQ68wCzGAn12F9UKcBEi3IR9n1fDADioiAr3Q4tAIYyjWIAUW2PYDmbaQKKY5QCiRaBK0wOEETyIQiyfDAH1I8iQHWPIwGqUgaKU19VPUzSmDUvASXEaBEAk3SNNIOg8BeF44QfRAQnQCzNPAyCpGQf44BQlS1MsoQIKcJtoF0Zy-M0oQ8kQ8knHWTTmzsMK9NIIRNnWDACl2YjTOCJLLJ-P9m3+U4unQF5djwTY0qIFiMPmLDtPQCrYr4Yge0ymAZBAZY+1q8Vx2TX4EABIEnD6tgBofeAKGqHAPyy4LGpAWb5vmBs+okdAVogBaeJNficsyCTtt2vt8UHGa5p2tbCWcdZWqIdrdk6vF9lEL8GzS7BNiy2LkVRZxZhIui8QJDEsWwXEQHxVF1k7MkKSg6lobBj6vtgZxRHWWZP2e0HYaXWA9HxtFd2mmM5PmBTMTDYmQHJCB1nWOgAAl1jwWZKlEUdxwchB-lTdKEQgdEjN2IhgbI18c0qeZ3uqUC3sh0F3wugBfLx0BzCWcQoHMsPDEM8DyL88FECdsAAcQKaoLG7aoiHWXYLCEYh1gsAAhT8nAwQKgbVoA}{Link}
\end{itemize}