### 2024-08-21 Findings (Shared Folder)/Chapters/branchingMorphogenesis.tex

\chapter{Branching Morphogenesis}

\begin{itemize}[nosep]
	\item Modified Fisher-KPP model for branching morphogenesis in VisualPDE (we have taken into account the segment): \href{https://visualpde.com/sim/?options=N4IgxiBcIG4JYCc4BM4GcQBoTIKq6hABEscA1MwgBlOQHU7raNoBGKj0sA6G7Aa0JEABAF5hrAHRVhcAHbCA2q0zsAugG4EAfQBGY4VWntZCxVVWa52meIDM000osBWTToCmBo6xdPzlhoANt7SMvLOmAAsmgC2oRz+FjEapLEAhtSSdgDsaXBM2AAOhADCAK5oAC4A9rGkRQg1uh4AYuVyhOUNTS0AMh5yAOZVABZQdhwWIAg8M3oAVOULABSsALQr5QDUMACUAPQr1lR7e8LrwrFLpAiU0J5LwtvCKzq6Byd7S6s7+xdBBYwW6MXikNAlaB2UjAyCKUDpBBNADupRqQRq5QQUAAbC4XHYXNhESiiIM0HAqgBPKBGIkgEk1ZEDYZjACy6QAHoRWKRGciAMpgdJBDyEdLlWp8pFMgAahD62k5BwATBdhNziTLkQBNBXaKmq9U04mSmoCjxVNEYrEAJXSwzFkAAZiK0B5sMLYh4EOkACqjS2ZSCTT3pb2+gAKowKIemXp96QAWjU6rTJAAOT3ozEIDKQ2CIFDoLg5rG6RFQV1Bd3ZuS1LEsau18A1eu56252l1hsIACiRQpGM6kCMVBUrfbWIAcuV6pB6WBKrVYgKsa6wE7mx6QB5Yroamgm26d3uD0eAIL7uCDKrp+lnw9oIhwZ3OypOow47CPo-R2NSC4ORUFmu77k+AoxnINKQOwP7gUeAqxKmYxyB4R7piq8HnmgApFB4S5BJWo6SFhYE4QGQbpjkGYuFE2DOkEcBFPhyCdli+ZVie2BMWhdAoOy5RBFAE4ZJy7F3CK5ROiARi5PkcgSWQUkyXwIA1DAPpETB27YBpWnpFSEndupmkINpA5Dm26YcPS+nmYZfaco0PKkPZ2l9PIHj8cggnCbBxQYlUfpUvhhBFERaE9M08hDFxNY7o0zRtB0XTRf0gwjOMcYcMUvQeCFYXQPIVQeEMvrCXlyXymCVUtHqtUgGg67pJu7QjrJpBjAgHgeEQAASHhwEMoxVEKIqfpIOTYJpYC1AgrQ3kEyDxS2yKjOkwU1JGQVpdgcjhjJvIAL6YAi2rGZAeIEvS-JknIFLUveWooiyWUctybDSii42iuKZrfXK+rKmqlyagy2oNSAiqGqDwgmgyZoWlaZYIPajqrTuCa+pRVTBqG4CHVGMYTPGRPJqm85GKBYCo5x0DwEgqAYNmNoIBW2Iutxk69seCU9h2qMmbTU79oOcDDjZE4i72s7zouy51GuCAblu3O-nzLYa1eug3vWz3kU+L5vh+6bfobf4k7BkhASB2EQVBMFwRbuHITUqHoSwRhkRreEEUJxHe-bR648GRg0XRDFMSxHhsXT6SQrpIC8d5AmjGyQkidgYlKSpWTydn8i50E0mFKZBk6dz7mGZd0zV1SlkS9ZJG2XpZkWc5nMgLybcV55fFpxn-kThF7uFTJEUOmKdWxZjdUpR13Tz29Yyk7lIBJS04+ECVZUVelHg1Z189Q2pzUq61C9l91vUDUNI1jcKf0kdNsAEfNi0eMtc8gOtm1+ttXaDNSAHW9FAFUx01DYGRCtaAS9f4FngRSWIfpKTPxAItNAgYEDrAANKRkjMINkNRkBf2EM6GoCBhAACFfRyDAFBIYRDKFFFGDUIYgx0IlmOkAA}{Link}.
	\item Guo pattern formation model for side branching: \href{https://visualpde.com/sim/?options=N4IgRgTiBcIAwDoCsIA0IDGMQDcCWEeAJngM5ohECqV2AIgPoCCFRAam-QwJKsDqfLgGVWAFxiI4ARnQZasAEpMAcnQBU3BkLVS4FDJ1hSKAay5MABAF4LkgEwW8AOwsBtOKikBdANyNu1rYIdgBsji7unr6MQoGSYc5uHt4+GHEIcHAOiZEpALYArulSCRHJvhAAFgD2zOlwAMzhSVE+TkU2kgAszbkVNTz1mY295akMcPXZZa0A5gCGeXnzU6OtAKYADqR4ADbVLp0IABxIaynrDFLFkzljREM9d60AZoG6574Uy9iIDQDs3zwvwoTmwXQom2wAGECqRRNU8pCINUwOsAGIFMGwFjoTYotEAGXWTlmokqMAaw3QEHkmDUTAAenY1EIAPTcAC0hQZAGoqrUmGoAJoUCCGelMlmxTkWdoaCz8gbcEViwSwDATCyyhZLeas7UWLY7fZOEWssUARWwRAZhsuUjUFmFiosAAphcyAJRst1SXkvEXeiikbEgSwBWKi9CkACOMFEEAK63QOBgrlA8wgKIA7tDqvsClBoCEkEgGkh0Fnc3QSTtRABPCTIKvZ6o54mk8kAWXmAA9sMZW7mhBh5rt1th5gUERRq+2ABrYQkMPtshyygfD9ui2Arhvrw1Nqsz6pCdaifOFiAKeakyfQF7j0gpzCLdYQeYAFUqF-mlI8N88g-eYAAVKmBaAqVkd9PwALWqRFm2OWQC2qItlihWB8EIEhyFQ68wCzGAn12F9UKcBEi3IR9n1fDADioiAr3Q4tAIYyjWIAUW2PYDmbaQKKY5QCiRaBK0wOEETyIQiyfDAH1I8iQHWPIwGqUgaKU19VPUzSmDUvASXEaBEAk3SNNIOg8BeF44QfRAQnQCzNPAyCpGQf44BQlS1MsoQIKcJtoF0Zy-M0oQ8kQ8knHWTTmzsMK9NIIRNnWDACl2YjTOCJLLJ-P9m3+U4unQF5djwTY0qIFiMPmLDtPQCrYr4Yge0ymAZBAZY+1q8Vx2TX4EABIEnD6tgBofeAKGqHAPyy4LGpAWb5vmBs+okdAVogBaeJNficsyCTtt2vt8UHGa5p2tbCWcdZWqIdrdk6vF9lEL8GzS7BNiy2LkVRZxZhIui8QJDEsWwXEQHxVF1k7MkKSg6lobBj6vtgZxRHWWZP2e0HYaXWA9HxtFd2mmM5PmBTMTDYmQHJCB1nWOgAAl1jwWZKlEUdxwchB-lTdKEQgdEjN2IhgbI18c0qeZ3uqUC3sh0F3wugBfLx0BzCWcQoHMsPDEM8DyL88FECdsAAcQKaoLG7aoiHWXYLCEYh1gsAAhT8nAwQKgbVoA}{Link}
	\item Guo pattern formation with better parameters \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oQCYCWATjQMYAWANggC5lgCK19TNp251CABgB0AVjJUIUWABEA+gEFZeAkuUBJWeUraAyrK7RJYgIy5GCkACVVAOUUAqHcqOvLY12LKMWiCWZADWhCqqAARRALwxFgBMMTRwMQDaYthRlgC6ANwqOjHxUZKJYilpUZnZeYWeJQkSAJwAbFUZWTkFjE1lEmJiyVGpXXUFKEj9kpYdo9W1Pfl0zAD2ajODAMydNd31cNNxzWIALHtL9asbxScDQ0O7C+PLjMqV90mXBwUA5mgUOgtsMfhN8ggAA4AZxorDW1VKkgAHFIwcsEMpLFtsS99uCqCDzuj6gAzfo+EkFMjocQSbYAdhpNHEZCw0DOuDgABUYGchrhIYQAMJIaEcNYoMiQuhrMAIABiSHZIHUgtl8oAMgg4H8OMwYNtHrgGDAOHQkAgTXZGK5VAA9RKuIwAeh0AFopnaANQ3NSuACaZDoQVtDqdRii7qiR3cUV9610geDhhA70q0YBQLQzqjUShsPhcEDzuDfFgVDtecxllcUQD8aiAAoA46AJQupuWb2kwPtsjQlXRYqRoO4aEARzNFqtIAAbjB0qA0HRZQB3YVreFIUTQNpSKTbKS4FfrxQ62EcACeMEkx5Ap7Wa+1uv1AFk0AAPQghE+rp9GIwaDsIQaBIBKZCPmuAAahCason4usk0bfn+65BrA8FXkheY3ie4FrEYnCbtudD2BgfwIDApLAdCs5ASgCB0Gg3LMJwaCGlkaaAkxaAAArMCy0BGjYPHMQAWmskq3hIyI2Fuaw7ugQqwBwO5gGsAQKepK7UbR9EIhKO7QnprB0fJcBGXQJGKbuXGMIZtkAKIwnCCIyVYFlWU4qAwPejBihKKBGDuNGMFR0A0WZs4ICgGnQiZkX6bgsXxdCqhxTQOpmHeKVxWsCWKDQpKkmKEWSG0eVpQJQmWNIDJiHJICpQV0JGIJcA3tAPhVa1RgoFJ+qIAlMmJL1CVGJCCABawunmBIY3NflCWsexMkMqinIgKSrA0JCU1UDZSloCpUXmSAu2IOQNBUO+SCsDA1ggOgn5HSGwGWnSjLMnAb14B9EUgP4uBrHOTGzV1Z2zqD4NoFeb23iDYN0BDLmFu581DPeMMo3DTmfjKP5kDjEOaqkCDXbdzBvvdj2CvCHDcleU2EJCs2INKGqpH8pnnTKcqKsqhBqiA-NajqeoGsJxqixqCBMyzsCpBwCB-MxD3qgLsGwMDssCxhQMDqFaDhUqKq6-qdAIAgigABIIDQfzMBwgHAeVEgMrgYOMBKdAKllrBULzs5rswaCM2sfEM8LbI8UTAC+uS4GuQewCLa4qaqA40Cg3I0BwIGwAA4kgaxRG+axUAgrBREYN0IFEABCzFwCw3MgPHQA}{Link}
	\item Guo mostly side branching: \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oQCYCWATjQMYAWANggC5lgCK19TNp251CABgB0AVjJUIUWABEA+gEFZeAkuUBJWeUraAyrK7RJYgIy5GCkACVVAOUUAqHcqOvLY12LKMWiCWZADWhCqqAARRALwxFgBMljE0cDEA2mLYUZYAugDcKjox8VGSllKp6VFZOflFnqUJEgDMAJwAbNWZ2bmFjM3lEmJiiT21fQ0oSEMV3VFpvfWFdMwA9mpzI60Tdf0FcLNxLWIALHtTqxu626Oju4s1+w2MymJ340-LBwDmaCh0J9LisCggAA4AZxorHWNTKkgAHFVvpNQQhlCkTsMsUs0QcqHdziCDgAzIY+EkNMjocRtADsNJo4jIWGgZ1wcAAKjAzqNcODCABhJCQjjrFBkcF0dZgBAAMSQbJA6gFMrlABkEHBfhxmDBWvdcAwYBw6EgEMa7IxXKoAHqJVxGAD0OgAtDNbQBqNabVSuACaZDoQRt9sdRiibqiR3cUR9Nx0geDhhAbw+0f+gLQTqjUQh0NhcEDTuDfFgVFteYxllcUQD8aiAAoAw6AJTOpuWL2kwPtsiQwWwSStAfK6IlSNB3CQgCOpvNlpAADcYBlQGg6DKAO5C9awpCiaCdKRSVpSXCbneKbXQjgATxgkgvICv623Wp1eoAsmgAB6ECEl5bu+RiMGg7CEGgSDimQb7bgAGoQGrKH+zrjNGAHATuQawCh97oXmj6XjB6xGJwe4HnQ9gYL8CAwKSEGQku4EoAgdBoFyzCcGgBrZGmALsWgAAKzDMtAho2IJHEAFrrBKT4SIiNj7ush7oEOIAcIeYDrAEqk6ZuDFMSxcLioekLGawzEqXA5l0JRalHvxjBmU5ACiUIwnCilWLZ9lOKgMAvoworiigRiHoxjD0dAjHWUuCAoLpkKWXFJm4ElKWQqoyU0NqZjPplyXrKlig0KSpKirFkidMV2WieJljSPSYjKSAWWlZCRhiXAj7QD49VdUYKDyXqiCpYpiRDalRjgggoWsEZ5gSNNHUlalXE8Yp9LIhyICkqwNDgvNVCOepaBDvFNkgEdiDkDQVA-kgrAwNYIDoH+50hhBFp0q0jK4CgaTfXgv2xSA-i4Osy7sUt-XXUuMNw2g97fU+0Ow3Q8OeYWPkraML7I9jqPuX+0qAWQxPwxqaQIA9T3MN+L1vQKsIcFy97zYQ4JLYgUrqmkvxWTd0qygqSqEKqIBi5q2q6vqElGjL6oIJz3OwGkHAIL8HGvWq4tIcOAvi7hkMDlFaAxYqypQ1pzB0AgCCKAAEggNC-MwHBgRBNUSPSuCw4w4p0PK+WsFQItLtuzBoBz6zCezUusoJlMAL55Lg26R7A04gNumnS9CKBcjQHCQbAADiSDrFE37rFQCCsFERiPQgUQAEIcXALBCyAadAA}{Link}.
	\item Guo symmetric side branch \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oQCYCWATjQMYAWANggC5lgCK19TNp251CABgB0AVjJUIUWABEA+gEFZeAkuUBJWeUraAyrK7RJYgIy5GCkACVVAOUUAqHcqOvLY12LKMWiCSlgGGIKG4ANaEKqoABPEAvIkWAEyJNHCJANpi2PGWALoA3Co6iSnxkmkAbJnZ8XkFxWWelakSYvXxWbn5haWMHdVdYhm9jc2DJShIIyE9fU0DrXTMAPZqC10AzA39LaVw88mdYgAsBytHJetbFWejYi+W19OtjMpiO+Pvq6UAOZoFDoX4TZYfUoIAAOAGcaKwNo0qjV-rcEMo3k8Qj9JocZlRfpd0TMAGYjHyk1pkdDiCS7ADstJo4jIWGgF1wcAAKjALi9cDDCABhJBwjgbFBkGF0DZgBAAMSQHJA6iFcoVABkEHBARxmDBdi98iAGDAOHQkAhcHQ7IxXKoAHppVxGAD0OgAtHNHQBqe5qVwATTIdCCDudrqM8S98RO7niAc2uhDYfCXx+ceBoLQbtj8VhCKRcBDbrDfFgVEdBcxllc8WDSfiAApgy6AJTuluWP1kkOdshw1UJCox0O4OEARwtVptIAAbjAcqA0HQ5QB3EUbJFIUTQWpSKS7KS4NebxS6hEcACeMEkp5A542G51eoNAFk0AAPQiRJ-ri+RiMGg7CEGgSCSmQz4bgAGoQWrKN+7oZHGv5noBG6hrAiE3ihBZ3mekEbEYnDbrudD2BggIIDAZKgXC84gSgCB0GgPLMJwaBGqazGsWgAAKzCstAxo2CC-EAFobFK94SAAHDYO4bHu6DCrAHB7mAGwBMpWlrnRDFMcikp7nChmsIxSlwKZdDkSp+68SZDkAKLwoiyJyVY1m2U4qAwI+jDipKKBGHu9GMLR0D0ZZ84ICg2lwuZ0VGbg8WJXCqgJTQupmA+aUJRsSWKDQZJkuKUWSLUBUZUJImWNIjJiIpIDpUVcJGMJcB3tAPg1e1RgoDJBqIElclpP1SVGDCCBBawBnmBIE2tYVSUcVxcmMvJUhciAZKsDQMIzVQ9mqWg6kxVZIAHYg5A0FQn5IKwMDWCA6Dfqd4agda9JMiycCfXg31RcEZAbAurHzT1l3zuDkNoDen33rgcN0FDbnFp5i0vI+qPo9+sp-mDENowjWpZAgd0PcwH5PS9QpIhwPI3jNhAwvNiAypqWSAhZV2yvKSoqoQ6ogAL2q6vqhqiSaGqC8zrOwFkHAIICbHPXLCrwbA-iawg2Gg5O4VoJFyqqrrIAGnQCAIIoAASCA0ICzAcMBoGVRIjK4BDjCSnQio5awVB8-OG7MGgTMbAJjMi+yElEwAvkUuAbsHsATiAG7qWqQ40CgPI0BwYGwAA4kgGzxB+GxUAgrDxEY90IPEABCbFwCwPMgAnQA}{Link}.
	\item Guo pure tip splitting \href{https://visualpde.com/sim/?options=N4IgRgqiBcIHYFMCuBbAhnOIA04BqM8y6mO4A6oYqhlrmAIpXG1lgBOhADAHQCsZACYQosACIB9AIJC8BcRICSQ8pQUBlIQBcYvLgEZcAY1Eg9BskflmeF42pD6yAa0KSpAAg8BeL3oBMXgCWcF4A2lzYHvoAugDckopevh68-gBswaEeEVGxCRLqyX62mR4h4ZHR8UbFqbZcgeXZudVxKEh1vPplFTlV+ewAFgD20l22AMxZlXnxcJ0+JVwALDP9c3HDY0lL9VwH+uut+UYSXBONxwPxAOZoKOiXTX0n8QgADgDOQQA2I9kUvpbNdNggJEc9t0Ls1Zm1BJcuAAOUFtABmdX0MNeNziZHQ3B4kwA7PigtwyFhoCtcHAACowFYHXAfQgAYSQXy0IxQZA+7BGYAQADEkFSQDIWQKhQAZBBwW5aIYwSYHSIgdjk6BadhIBC4dimIwAKikAD1-Mb1AB6RQAWg6poA1NtpMaAJpkdjWE3my1FO0eBbGpIu0ZKD1ehxnC6B+6PNBWjyBz4-f5wD1Wr1MWCCU3Jjzg-TGjzujxOjwACndFoAlNbK-onWiPXWyF9WbBeCt2+LPEkip7cF8AI4wHV63AANxgYVAaHYAoA7myRv8kJxoOk+HxJnxcAvl2J5T8tABPXT8A+LkZLuUKpUAWTQAA9CE5r8v1EY0L8EIQ0CQbkyEPW8AA1CBlCQX2tQJAzfT9b09WAoLPWCCwvA8gJGdQEC0Vd13YAAlDBbn-aA0V-L59RAH8UAQdg0DpIY8LQFV1Tohi0AABSGLVVWMB4uIALRGHlLyRYw1xGDd0E7EAtA3MARksaSlIXGBKN+aipLgbkNy+TSqJoowAX09gCJkzcOLMqyAFFvj+AFLwMXTzIAOVQGB91ozluRQdQN0ooxyK0nSQAQFBlK+QyKOM3BIuir4pCioJ5R0aBeB8xKRhisQgjRNFOXI3h0gSqLcq+XitWBPhiWRcqkvUPi4AvaAsUayr1BQMSlUQGLL38TqYvUD4ECMJBfg0zKeCGiKKpi5jWMvYkkT4GkQDRX4gg+MbBEs2S0E7MKaO2xByCCQQn0mmBDBAdAXwO71fz1QkSTJOAnrwF7yLMMgRinBiprak7cABoG0DPJ7dDBwH2GBhy02cmaDh88H4chuyX35d9-rh4GZRCBALquoZHxu9qWX+LQ6TPMbCA+KbED5aUQluIztJo-lBRFMVCElEBudleVFWVaBVWZQXpQQWn6dgEItAQW5GN+FmeYgrs1aFZC-uHIK0BC0VxS4MglXYBAEDEAAJBAgluIYtG-X8Sp4Ylp3G7l2GFNLfkEDnwqXIY0BpkZuOp-nKSE3GAF8YlwJc-dgIcQCXeSBZ+FA6SCLQ-0IABxJARg8R8RkEBBfg8dRLoQDwACFGLgIwWvZ6OgA}{Link}.
\end{itemize}